Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe

### 2017-08-21 manuscript/usses_water_ms.tex
updates regression figure with lines trimmed to data ranges

Ecosystem response to precipitation extremes depends on the
physiological responses of constituent species and the rate at which
community composition shifts to favor species better able to take
advantage of, or cope with, new resource levels \citep{Smith2009}.
Previous work has shown that community compositional shifts can be both
rapid, on the order of years \citep{Hoover2014}, and slow, on order of
decades \citep{Knapp2012, Wilcox2016}. Thus, a lingering question is how
the time scales of ecosystem response and community change differ among
ecosystems, which can be answered by manipulating precipitation to reach
extreme levels. The results of our five year experiment in a sagebrush
steppe conform to the first of our four predictions: neither ecosystem
functional response nor community composition changed with chronic
alteration in water availability (Fig. 1, top left).

The similarity of ecosystem functional response (Fig. 3) and community
composition (Fig. 5) across treatments is surprising because grasslands
generally, and sagebrush steppe specifically, are considered
water-limited systems. Indeed, we expected ecosystem functional
response, community composition, or both to change with precipitation
treatment, landing us in any box of Fig. 1 \emph{except} the top left.
So, why did our treatments fail to induce ecosystem or community
responses? We can think of three reasons; two are limitations of our
study, and one is the life history traits of the species in our focal
communities. We first discuss the potential limitations of our study,
and then discuss the biological explanation.

First, it could be that our manipulations were not large enough to
induce a response. That is, maybe a 50\% decrease/increase in any given
year is not abnormal given our site's historical range of variability
\citep{Knapp2017}. We cannot definitively rule out this possibility, but
we have reason to believe our manipulations \emph{should} have been
large enough. Using the methods described by \citet{Lemoine2016}, we
calculated the percent reduction and increase of mean growing season
precipitation necessary to reach the 1\% and 99\% extremes of the
historical precipitation regime at our site. The 1\% quantile of
precipitation at our site is 110 mm, a 47\% reduction from the mean, and
the 99\% quantile is 414 mm, a 77\% increase from mean growing season
precipitation (Appendix 3). Thus, our drought treatment represented
extreme precipitation amounts, especially in years where ambient
precipitation was below average (Fig. 2A). The irrigation treatment may
have been too small, however.

>>>>

Ecosystem response to precipitation extremes depends on the
physiological responses of constituent species and the rate at which
community composition shifts to favor species better able to take
advantage of, or cope with, new resource levels \citep{Smith2009}.
Previous work has shown that community compositional shifts can be both
rapid, on the order of years \citep{Hoover2014}, and slow, on order of
decades \citep{Knapp2012, Wilcox2016}. Thus, a lingering question is how
the time scales of ecosystem response and community change differ among
ecosystems, which can be answered by manipulating precipitation to reach
extreme levels.

The results of our five year experiment in a sagebrush steppe conform to
two of our four predictions, depending on treatment. Under chronic
drought, neither ecosystem functional response nor community composition
changed (Fig. 1, top left). But under irrigation, ecosystem functional
response was different from the control treatment while community
composition remained unchanged (Fig. 1, top right). Thus, we found some
evidence for a saturating response at high water availability (Fig. 3),
but the response of ANPP to soil moisture was weak overall (Table 1).

The similarity of ecosystem functional response (Fig. 3) and community
composition (Fig. 5) between drought and control treatments is
surprising because grasslands generally, and sagebrush steppe
specifically, are considered water-limited systems. Indeed, we expected
ecosystem functional response, community composition, or both to change
under the drought treatment, landing us in any box of Fig. 1
\emph{except} the top left. So, why did our drought treatment fail to
induce ecosystem or community responses? We can think of three reasons;
two are limitations of our study, and one is the life history traits of
the species in our focal communities. We first discuss the potential
limitations of our study, and then discuss the biological explanation.

First, it could be that our manipulations were not large enough to
induce a response. That is, maybe a 50\% decrease/increase in any given
year is not abnormal given our site's historical range of variability
\citep{Knapp2017}. We cannot definitively rule out this possibility, but
we have reason to believe our manipulations \emph{should} have been
large enough. Using the methods described by \citet{Lemoine2016}, we
calculated the percent reduction and increase of mean growing season
precipitation necessary to reach the 1\% and 99\% extremes of the
historical precipitation regime at our site. The 1\% quantile of
precipitation at our site is 110 mm, a 47\% reduction from the mean, and
the 99\% quantile is 414 mm, a 77\% increase from mean growing season
precipitation (Appendix 3). Thus, our drought treatment represented
extreme precipitation amounts, especially in years where ambient
precipitation was below average (Fig. 2A). The irrigation treatment may
have been too small, yet that is the treatment where we observed an
effect (Fig. 3).
